---
title: "CSPackage for FARS Data"
author: "Christopher Sumter"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CSPackage for FARS Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  root.dir = system.file("testdata",package = "CSPackage")
)
library(CSPackage)

```

The CSPackage includes some functions for importing, exploring and visualising the US National Highway Traffic Safety Administration's ([Fatality Analysis Reporting System](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars)), which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes.

## Importing FARS data for a given year

FARS data has a specific file name convention therefore to read *all* fields for a single year, wrap the desired year with **make_filename()** and call **fars_read()**

```{r readYears}

singleYearsData <- fars_read(filename = make_filename(2015))

head(singleYearsData, n=10)

```

## Summarize number of fatalities by state and year

The CSPackage also includes a useful summary function to show the number of fatalities across the United States for a given number of years.

```{r summariseFARS }
fars_summarize_years(years = 2013:2015)

```

##Visualizing Fatalities

In terms of visualization one can create a simple map plot of fatalities for a given year and state (identified by its ID) using the **fars_map_state()** function.

```{r visualizeFatalities}

fars_map_state(year = 2014, state.num = 13)

```


